---
title: "Visualization of surface marker data"
author: "Pang Wei Koh"
date: "5/24/2016"
output: html_document
---

The script reads in values from surface_markers.csv, a CSV where each column is a cell type and each row is a surface marker. 
For convenience, you can download this file from http://mitra.stanford.edu/kundaje/pangwei/mesoderm_data/surface_markers.csv.

We first load in the data, filtering out all surface markers that are homogeneously expressed (or not expressed) across
all cell types. We use an arbitrary threshold of 30% for the purposes of data visualization.

```{r warning=FALSE, message=FALSE}
library(RColorBrewer)
library(gplots)

markerData <- read.csv('surface_markers.csv', header = TRUE)
rownames(markerData) <- markerData[, 1]
markerData <- markerData[, -1]
colnames(markerData) <- c("D0 hESC", "D1 APS", "D2 PXM", "D3 Somite", "D5 Dermo", "D6 Sclero", "D3 Cardiac")

threshold <- 30
maxMarker <- apply(markerData, 1, max)
minMarker <- apply(markerData, 1, min)
idxToKeep <- (minMarker < 100-threshold) & (maxMarker > threshold)
```

We then use the remaining surface markers to construct a heat map, keeping the original order of the cell types:
```{r warning=FALSE, message=FALSE}

hmcol <- colorRampPalette(brewer.pal(9, "YlGn"))(100)
markerHeatmap <- heatmap.2(as.matrix(markerData[idxToKeep, ]),
          col = hmcol,
          Colv = FALSE,
          dendrogram = 'none',
          trace = 'none',
          key = TRUE,
          density.info = 'none',
          margin = c(12, 10)
)
```

These are the row names (names of surface markers), in order from top to bottom:
```{r}
rev(rownames(markerData[idxToKeep, ])[markerHeatmap[['rowInd']]])
```

```{r}
sessionInfo()
```